Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2014-12-18 parts/example.tex
revise TCRSIG example + justifictions of our method over [13]

The Boolean model has been automatically encoded into a Process Hitting with 2 classes of priority%
\footnote{Files are available at
\url{http://maxime.folschette.name/underapprox-tcrsig94.zip}.}.
Then, we verified the reachability for the independent activation of 4 outputs of the signalling
cascade (SRE, AP1, NFkB, NFAT) from all possible input combinations (CD4, CD28, TCRlig) using our
new reachability under-approximation (answering either \emph{yes} or \emph{inconclusive}) and a 
previously defined reachability over-approximation \cite{PMR12-MSCS} (answering either \emph{no} or
\emph{inconclusive}).
All result in conclusive decisions, and the under-approximation has been satisfied in 12 cases (over
32) proving the satisfiability of the concerned reachability property in the encoded Boolean network
(and non-satisfiability in the other cases).

Computations times are in the order of a few hundredths of a second on a 2.4GHz processor with 2GB
of RAM.
To give a comparison, we did the same experiments with a standard symbolic model-checker, libDDD
\cite{libddd}, known for its good performances, the input model being the Boolean network expressed
as a Petri net.
However, due to the large scale of the model, the program runs out of memory for all the experiments.

>>>>

The Boolean model has been automatically encoded into a Process Hitting with 2 classes of priority%
\footnote{Files are available at
\url{http://maxime.folschette.name/underapprox-tcrsig94.zip}.}
\modMF{
by using a translation equivalent to the one given in \pref{sec:dn}.
\todo{?}
}
Then, we verified the reachability for the independent activation of 4 outputs of the signalling
cascade (SRE, AP1, NFkB, NFAT) from all possible input combinations (CD4, CD28, TCRlig) using our
new reachability under-approximation (answering either \emph{yes} or \emph{inconclusive}) and a 
previously defined reachability over-approximation \cite{PMR12-MSCS} (answering either \emph{no} or
\emph{inconclusive}).
All result in conclusive decisions, and the under-approximation has been satisfied in 12 cases (over
32) proving the satisfiability of the concerned reachability property in the encoded Boolean network
(and non-satisfiability in the other cases).

\modMF{
By analysing the detail of the results, one can find out that one of the output, NFkB, can never be activated.
In other words, FkB$_1$ is never reachable, whatever the state of the inputs,
while some combinations the other outputs (SRE, AP1, NFAT) can be activated for some configurations of inputs.
We thus want to check if these three outputs can be activated together,
that is, if the activation of one of the outputs does not prevent the activation of one of the others.
For this, as proposed in \pref{ssec:simult-ua},
we add a new cooperative sort $\total$ between SRE, AP1 and NFAT to the model,
and a sort $\reach$ with two processes ($\PHl_\reach = \{ \reach_0, \reach_1 \}$)
with the action $\PHhit{\total_{111}}{\reach_0}{\reach_1}$.
Finally, checking the reachability of $\reach_1$ in all configurations of the inputs
is always conclusive,
and the existence of positive answers allows to conclude that
it is possible to reach a state where SRE, AP1 and NFAT are simultaneously active.
}

We held all these experiments on a laptop computer with a 2GHz processor
and 1GB of RAM.
\modMF{
Computations times were in the order of a few hundredths to tenths of a second
in each case (reachability of $\reach_1$ included).
}
To give a comparison, we did the same experiments with a standard symbolic model-checker, libDDD
\cite{libddd}, known for its good performances, the input model being the Boolean network expressed
as a Petri net.
However, due to the large scale of the model, the program runs out of memory for all the experiments.
\modMF{
This limitation shows that studying the dynamics of models is strongly limited to the size of the model for standards model checkers,
while our method is able to conclude with limited memory and computation time usage.
Furthermore, given that the complexity of the static analysis presented in this paper
is polynomial in the size of the models,
it is expected that our work will be scalable to models that are larger,
even by orders of magnitude.
%---~that is, containing thousands of components.
}

\modMF{
Finally, we note that similar experiments were conducted in~\cite{PMR12-MSCS}.
However, if these experiments allowed to obtain some information
on these model to some extent,
it did not provide a formal proof of the reachabilities that have been
checked in the example above.
Indeed, without any priorities in the update of the cooperative sorts,
some unwanted behaviours were possible,
such as those described in \pref{ssec:priorities}.
Only the method presented in this paper provides a formal proof that
the observed behaviours are the result of the exact dynamics of the system,
which was originally described in a format equivalent to synchronous Discrete Networks.
}

### 2014-12-18 parts/sa.tex
revise TCRSIG example + justifictions of our method over [13]

\begin{example}
  Let $\PH' = (\PHs, \PHl, \PHh'^{\langle 1 \rangle})$ be the “merged” version of the PH
  depicted in \pref{fig:ph-livelock}, as it was already defined in \pref{ssec:priorties},
  that is: $\PHh'^{\langle 1 \rangle} = \PHh^{(1)} \cup \PHh^{(2)}$.
  In other words,
  $\PH'$ is how $\PH$ would have been represented in the semantics without priorities.
  We showed in \pref{ssec:priorties} that,
  due to spurious behaviours inherent to the cooperative sorts in this semantics,
  process $c_1$ is reachable
  from the initial context $\ctx = \PHstate{a_1, b_0, c_0, ab_{10}}$.
%   This is confirmed by the original under-approximation
%   developed in~\cite{PMR12-MSCS}, which gives the same conclusion.

@@

\subsection{Reachability of a state}

>>>>

\begin{example}
  Let $\PH' = (\PHs, \PHl, \PHh'^{\langle 1 \rangle})$ be the “merged” version of the PH
  depicted in \pref{fig:ph-livelock}, as it was already defined in \pref{ssec:priorities},
  that is: $\PHh'^{\langle 1 \rangle} = \PHh^{(1)} \cup \PHh^{(2)}$.
  In other words,
  $\PH'$ is how $\PH$ would have been represented in the semantics without priorities.
  We showed in \pref{ssec:priorities} that,
  due to spurious behaviours inherent to the cooperative sorts in this semantics,
  process $c_1$ is reachable
  from the initial context $\ctx = \PHstate{a_1, b_0, c_0, ab_{10}}$.
%   This is confirmed by the original under-approximation
%   developed in~\cite{PMR12-MSCS}, which gives the same conclusion.

@@

\subsection{Reachability of a state}
\label{ssec:simult-ua}
